Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-05-05 text/Manuscript.Rmd
Fig6 update

cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Veraszto2018]. In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS). The fluorescence intensity of DAF-FM, which specifically detected INNOS, showed a transient tonic-like increase dependent on UV stimulation of the cPRCs. The graph shows a decrease in intensity after UV light is switched off, but the response of DAF-FM to NO is irreversible and this decrease is due to breaching by the relatively strong excitation light. Attempts to measure changes in the intensity of the calcium indicator RGECO1a simultaneously with DAF-FM imaging did not lead to detection of changes in fluorescence intensity compared to GCaMP6s and could only visualise the cPRCs. In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs. This tonic response may involve dopaminergic nerves based on single-cell data (Figure 6A) [@puopolo2001][@zhang2007][@Randel2014a]. In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018]. These suggest that UV stimulation of cPRCs causes the production of NO via a depolarising response of INNOS.

@@

![**Figure 6. Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq. The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells. The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene. **(B)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}

@@

Whole-mount immunostaining of 2 day old Platynereis larvae fixed with 4% paraformaldehyde were carried out using primary antibodies raised against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide neuropeptide in rabbit, plus a commercial antibody raised against acetylated tubulin in mouse (Sigma T7451). The synthetic peptides contained an N-terminal Cys that was used for coupling during purification. Antibodies were affinity purified from sera as previously described [@Conzelmann2012]. Immunostainings were carried out as previously described [@Conzelmann2012]. The NOS promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato. Larvae injected with promoter constructs (ca. 250 ng/ml) were analysed for reporter expression at 3 days post fertilization using an AxioImager Z.1 fluorescence wide-field microscope (Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings. The protocol followed for immunostaining of HA-tagged reporters was recently described [@Veraszto2017]. Specimens were imaged with a LSM 780 NLO Confocal Microscope (Zeiss, Jena).

@@

Green cGull was used for the cGMP assay [@Matsuda2017]. A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a Platynereis dumerilii cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence. Cos-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression. This cell line was purchased from Angio-proteomie (CAT no. cAP-0203). The Cos-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat. No. 10566016) supplemented with 10% fetal bovine serum (Thermo; Cat. No. 10082147). Upon reaching confluency of approximately 85%, we transfected the cells with the plasmid containing Green cGull-T2A-NITGC1. Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001). Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium. For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2 , 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose). Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = xx). Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (????). S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St. Louis, MO, USA) . The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss). Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition. Imaging data analysis was performed using ImageJ (National Institutes of Health, Bethesda, MD, USA).

>>>>

![**Figure 6. Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq. The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells. The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene. **(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}

cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Veraszto2018]. In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS). The fluorescence intensity of DAF-FM, which specifically detected INNOS, showed a transient tonic-like increase dependent on UV stimulation of the cPRCs. The graph shows a decrease in intensity after UV light is switched off, but the response of DAF-FM to NO is irreversible and this decrease is due to breaching by the relatively strong excitation light. Attempts to measure changes in the intensity of the calcium indicator RGECO1a simultaneously with DAF-FM imaging did not lead to detection of changes in fluorescence intensity compared to GCaMP6s and could only visualise the cPRCs. In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs. This tonic response may involve dopaminergic nerves based on single-cell data **(Figure 6)** [@puopolo2001][@zhang2007][@Randel2014a]. In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018]. These suggest that UV stimulation of cPRCs causes the production of NO via a depolarising response of INNOS.

@@

@@

Whole-mount immunostaining of 2 day old Platynereis larvae fixed with 4% paraformaldehyde were carried out using primary antibodies raised against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide neuropeptide in rabbit, plus a commercial antibody raised against acetylated tubulin in mouse (Sigma T7451). The synthetic peptides contained an N-terminal Cys that was used for coupling during purification. Antibodies were affinity purified from sera as previously described [@Conzelmann2012]. Immunostainings were carried out as previously described [@Conzelmann2012]. The NOS promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato. Larvae injected with promoter constructs (ca. 250 ng/ml) were analysed for reporter expression at 3 days post fertilization using an AxioImager Z.1 fluorescence wide-field microscope (Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings. The protocol followed for immunostaining of HA-tagged reporters was recently described [@Veraszto2017]. Specimens were imaged with a LSM 780 NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).

@@

Green cGull was used for the cGMP assay [@Matsuda2017]. A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence. Cos-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression. This cell line was purchased from Angio-proteomie (CAT no. cAP-0203). The Cos-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat. No. 10566016) supplemented with 10% fetal bovine serum (Thermo; Cat. No. 10082147). Upon reaching confluency of approximately 85%, we transfected the cells with the plasmid containing Green cGull-T2A-NITGC1. Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001). Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium. For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2 , 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose). Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = xx). Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (????). S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St. Louis, MO, USA) . The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss). Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition. Imaging data analysis was performed using ImageJ (National Institutes of Health, Bethesda, MD, USA).
